Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2017-01-05 manuscript.tex
latest

As we will see, a generalized version of the classic lottery model of \cite{chesson_1981} is well suited for this purpose. The lottery model gets its name from the way in which it represents competition: the growth of each type depends on how many territories it secures from a fixed territorial stock, which is determined by the number of contestants that it produces relative to the other types in the population weighted by a coefficient for each type representing competitive ability. This is much simpler than having coefficients for the pairwise effects of types on each other (e.g. the $\alpha$ coefficients in the generalized Lotka-Volterra equations), or than modeling resource consumption explicitly \citep{tilman_1982}. The classic lottery model  breaks down at low absolute abundances at the scale of territorial contests (section ``Mean field approximation''), but this was not important for its original application to reef fishes, where a huge number of larvae from each type compete to secure territories each generation \citep{chesson_1981}. Here we analytically extend the classic lottery model to correctly account for low density behavior.

>>>>

As we will see, a generalized version of the classic lottery model of \cite{chesson_1981} is well suited for this purpose. In the classic lottery model, mature individuals (henceforth ``adults'') each require their own territory, whereas newborn individuals (henceforth ``propagules'') disperse to, and subsequently compete for, the territories made available by the death of adults each generation. Territorial contest among propagules leaves a single victorious adult per territory, with the victor chosen at random from the propagules present (akin to a lottery) with probabilities weighted by a coefficient for each type representing competitive ability. This lottery competition is much simpler than having coefficients for the pairwise effects of types on each other (e.g. the $\alpha$ coefficients in the generalized Lotka-Volterra equations), or than modeling resource consumption explicitly \citep{tilman_1982}.

In addition, the population-genetic Wright-Fisher model (each generation is a sample of constant size, with replacement, from the previous one) can be thought of as a special case of the classic lottery model.

The classic lottery model assumed that the number of propagules and available territories is essentially infinitely large, where each available territory is swamped with competing propagules every generation (its original application was to reef fishes, where a huge number of larvae from each species compete to secure territories each generation; \citealt{chesson_1981}).

As a result, type abundances change deterministically (in a constant environment). Nevertheless,

which models gen can be interpreted as a type of lottery model

Adults can persist over multiple generations, and so the classic lottery model allows for overlapping generations;

this is essential for the coexistence mechanisms that

propagule dispersal and competition is resolved over one generation.

This is a feature

 breaks down at low absolute abundances at the scale of territorial contests (section ``Mean field approximation''),

Here we analytically extend the classic lottery model to correctly account for low density behavior.
